Dispersion of short waves riding on longer waves

### 2024-05-08 paper/paper.tex
WIP governing equations

\authors{M. Curcic\affil{1,2}}

@@

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

\begin{abstract}
Abstract goes here
\end{abstract}

\section*{Plain Language Summary}
Plain language summary goes here

\section{Introduction}

@@

Paragraph: Describe the conditions, assumptions, and limitations.

@@

The L-HS solution can be most readily obtained by starting from the dispersion
relationship for short waves:

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency,
$g$ is the gravitational acceleration,
and $U$ is the mean advective current in the direction of the wave propagation.
For simplicity, we will limit the derivation to gravity waves and ignore
surface tension effects, however, the solution can be as well derived for
gravity-capillary waves by carrying the surface tension terms.

Next, take the conservation of wave crests:

\begin{equation}
\label{eq:wave_crests1}
\dfrac{\partial k}{\partial t}
+ \dfrac{\partial \omega}{\partial x}
= 0
\end{equation}
where $\partial/\partial t$ and $\partial/\partial x$ are partial derivatives
in time and space, respectively.
Insert (\ref{eq:dispersion}) into (\ref{eq:wave_crests1}) to get:

\begin{equation}
\label{eq:wave_crests2}
\dfrac{\partial k}{\partial t}
+ C_g \dfrac{\partial k}{\partial x}
+ U \dfrac{\partial k}{\partial x}
+ k \dfrac{\partial U}{\partial x}
= 0
\end{equation}
where $C_g$ is the group velocity of short waves.
At this step, we assume that $g$ does not vary in space and time.
We will see later that to a second order $g$ of short waves varies in presence
of long waves.
If we follow the short waves in a Lagrangian frame, we can re-write
(\ref{eq:wave_crests2}) as:

\begin{equation}
\label{eq:wave_crests3}
\dfrac{d k}{dt}
+ k \dfrac{\partial U}{\partial x}
= 0
\end{equation}

>>>>

\authors{Milan Curcic\affil{1,2}}

@@

\begin{keypoints}
\item We revisit analytical and numerical solutions for hydrodynamic modulation of short waves by longer waves from linear wave theory.
\item We extend the original solution by \citeA{longuet1960changes} to a higher-order solution.
\item In contrast to earlier studies, we find that there exist stable numerical solutions for modulation by realistic long-wave groups.
\end{keypoints}

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that pertains to remote sensing, the
interpretation of in situ wave measurements, and numerical wave forecasting.
In this paper we revisit the early analytical work by \citeA{longuet1960changes},
as well as the more recent numerical simulations by \citeA{peureux2021unsteady},
to re-evaluate the changes in short-wave properties due to the presence of longer waves.
We present two distinct improvements to the current state of the art:
First, we derive a higher-order, non-linear solution for the short-wave wavenumber
and action density as function of the long-wave phase and steepness.
This solution suggests an increase in short-wave wavenumber by 18 and 38\% for
long-wave steepenesses of 0.1 and 0.2, respectively, compared to the original
linear solution.
It is also in close agreement with the full numerical solutions up to the
long-wave steepness of 0.25.
Second, we find that the full numerical solutions, which were previously found
to be unstable for homogeneous initial conditions, are in fact stable for more
realistic long-wave groups.
The key physical interpretation from these results is that the short waves can
indefinitely grow and steepen only in the presence of inifinite long-wave trains,
an unlikely occurence in the ocean.
\end{abstract}

\section*{Plain Language Summary}
The ocean surface is typically covered by waves of many wavelengths.
Longer waves bunch up and dilate the shorter waves, and this process is
important to consider for both the measurements and the forecasting of ocean
waves.
Although there have been analytical and numerical solutions for how the short
waves should change in the presence of longer waves, they have been simplified
and incomplete.
This paper revisits and extends these solutions to propose a more accurate
description of how the short waves should change in the presence of longer
waves.
\section{Introduction}

@@

In this paper we consider inviscid, irrotational, and incompressible flow.
In these conditions the linear (sine-like), deep-water, surface gravity waves
obey the dispersion relation:

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency, $g$ is the gravitational acceleration,
$k$ is the wavenumber, and $U$ is the mean advective current in the direction
of the wave propagation.
A current in the direction of the waves thus increases its absolution frequency,
wavenumber remaining unchanged.
An important limitation here is that $U$ must be slow-varying on the scales of
the wave period.
In the context of short waves riding on long waves, the mean advective current
is the horizontal near-surface orbital velocity of the long wave.

The second important relation is the conservation of wave crests:

\begin{equation}
\label{eq:wave_crests}
\dfrac{\partial k}{\partial t}
+ \dfrac{\partial \omega}{\partial x}
= 0
\end{equation}
where $t$ and $x$ are time and space, respectively.
This relation states that the wavenumber must change when the frequency varies
in space.

Equations (\ref{eq:dispersion}) and (\ref{eq:wave_crests}) only describe the
evolution of the wavenumber.
To describe the evolution of the wave amplitude, we use the conservation of wave
action:

\begin{equation}
\label{eq:wave_action}
\dfrac{\partial N}{\partial t}
+ \dfrac{\partial}{\partial x} \left[\left(C_g + U\right)N\right]
= 0
\end{equation}
where $N$ is the action density, and $C_g = \partial \omega / \partial k$ is the
group speed.
In deep water, the group speed is half the phase speed, so $C_g = 1/2\sqrt{g/k}$.

As the short waves ride on the surface of the long wave, defined by the elevation
$\eta_L = a_L cos{\psi}$, where $\psi = k_L x - \omega_L t$ is the long-wave phase,
they move in an accelerated reference frame and will thus experience effective
gravitational acceleration that changes in both space and time.
Inserting equation (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:

\begin{equation}
\label{eq:wavenumber}
\dfrac{\partial k}{\partial t}
+ \left(C_g + U\right) \dfrac{\partial k}{\partial x}
+ k \dfrac{\partial U}{\partial x}
+ \dfrac{1}{2} \sqrt{\dfrac{k}{g}} \dfrac{\partial g}{\partial x}
= 0
\end{equation}

Similarly, we expand the equation \ref{eq:wave_action} to obtain:

\begin{equation}
\label{eq:wave_action2}
\dfrac{\partial N}{\partial t}
+ \left(C_g + U\right) \dfrac{\partial N}{\partial x}
+ N \dfrac{\partial U}{\partial x}
+ N \dfrac{\partial C_g}{\partial x}
= 0
\end{equation}

@@
